\sysfull

### 2018-08-07 paper/asplos19/intro.tex
Final pass on section 4 (Carlo) and new figure 5
change memory manager to page privatizer

\item[C2] \emph{Dynamic Memory Consistency Handling:} merging static tasks on the fly rises  the need for dynamic memory consistency handling. This leads to the second challenge: how to dynamically detect inter-coalesced-task data dependencies and ensure efficient protection against power interrupts? \sys addresses this challenge by using a novel approach called \emph{dynamic page privatization}---performing real-time dependency tracking to enable protection on a task transition. Individual variables tracking, however, slows down a system dramatically. Therefore, \sys keeps memory consistent through \emph{memory virtualization} optimized for bulk accesses to task-shared data with high locality.

@@

The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
Section~\ref{sec:systemdescription} illustrates an overview of \sys. Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism. The dynamic page privatization is explained in Section~\ref{sec:memory_virtulaization}. Implementation details of \sys are given in Section~\ref{sec:implementation}. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results. Section~\ref{sec:related_work} positions \sys in the context of related work and Section~\ref{sec:conclusions} concludes and discusses future work.

>>>>

\item[C2] \emph{Dynamic Memory Consistency Handling:} merging static tasks on the fly rises  the need for dynamic memory consistency handling. This leads to the second challenge: how to dynamically detect inter-coalesced-task data dependencies and ensure efficient protection against power interrupts? \sys addresses this challenge by using a novel approach called \emph{page privatization}---performing real-time dependency tracking to enable protection on a task transition. Individual variables tracking, however, slows down a system dramatically. Therefore, \sys keeps memory consistent through \emph{memory virtualization} optimized for bulk accesses to task-shared data with high locality.

@@

The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
Section~\ref{sec:systemdescription} illustrates an overview of \sys. Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism. The page privatization is explained in Section~\ref{sec:memory_virtulaization}. Implementation details of \sys are given in Section~\ref{sec:implementation}. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results. Section~\ref{sec:related_work} positions \sys in the context of related work and Section~\ref{sec:conclusions} concludes and discusses future work.

### 2018-08-07 paper/asplos19/paging.tex
change memory manager to page privatizer

%\note{additional motivation for coala memory virtualization] we should emphasize that Coala programming model removes the burden of identifying the write-after-read dependencies from the programmer  as compared to Chain. And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}

@@

\sys's {\em virtual memory manager} (VMM) is a core enabler of \sys task adaptation mechanism. Its necessity is illustrated by example given in Figure~\ref{figure:coalescedWar}: resolving WAR dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent with task coalescing. Since task coalescing induces dynamic WAR dependencies,  a new memory consistency handling technique is mandatory so that these dependencies are resolved and handled at runtime.

VMM overview is given in Figure~\ref{figure:mem-man}.
VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
\texttt{\underline{private}} holds the consistent version of pages after each commit (and on reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
The VMM prohibits applications from directly accessing these buffers.
Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks.
When a coalesced task completes, the VMM ensures that {\em dirty} (i.e., modified) pages are committed to FRAM and visible to subsequent tasks.
If a power failure interrupts a task, all temporary modifications to pages in \texttt{working} (and \texttt{\underline{shadow}}) occurred after the last commit are discarded in order to keep data consistency.
% If the power interrupts a task, the VMM ensures the consistency of the data by relying on its \texttt{\underline{private}} buffer. The VMM ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
On a commit, the VMM atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
% SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM. A shadow page stores the data from an updated page that needs to be committed back to non-volatile memory, but that exceeded the size of the volatile working memory.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/mem-man.pdf}
    \caption{Interaction between task, memory manager and memory buffers upon \texttt{RP} and \texttt{WP}, and order of occurrence. Steps 2, 3, 4 and 5 are conditional.}
    \label{figure:mem-man}
\end{figure}

%\textbf{System Overview.} Figure~\ref{figure:mem-man} gives an overview of \sys's virtual memory management system. A page has a home location in non-volatile memory which we refer to as the page's \texttt{\underline{private}}. When a task accesses a page, the VMM swaps the page from its home location into the fully\hyp{}associative \texttt{working} buffer that is stored in volatile memory.  The \texttt{working} buffer is allocated to take up the volatile memory that remains after space for stack is allocated.

%The VMM utilizes a two-phase commit to write dirty pages from the \texttt{working} buffer back to their home location in \texttt{\underline{private}}. It optimizes this process by using an indirection table to abstract the location of the \texttt{\underline{private}} buffer and to eliminate the need for the second page copy. The VMM is forced to swap out a dirty page from the \texttt{working} buffer when a task requests a page that is not in the \texttt{working} buffer, and there is no free page in the working buffer (i.e., a page fault). On a page fault, the VMM moves the page to the direct\hyp{}mapped non-volatile \texttt{\underline{shadow}} buffer.

@@

After address translation (Figure~\ref{figure:mem-man}, Step 6), a task accesses the protected variable's location in the volatile \texttt{working} buffer (Step 7). The VMM keeps track of the page tags for the pages currently resident in the \texttt{working} buffer. When a task accesses a variable, it compares the variable's page tag to tags of the pages in {\tt working} (Algorithm~\ref{algo:rwar}, Line 2). If the accessed variable's page tag is not found in the page buffer, the operation incurs a {\em page fault} (Line 4). The byte is accessed in the page buffer at the index of the resident page and the variable's page offset (Lines 5--6).

@@

When accessing a protected variable with \texttt{RP} or \texttt{WP}, the memory manager first searches the variable's page in the \texttt{working} buffer (Figure~\ref{figure:mem-man}, Step 1).
If the page is not found there, a page fault is incurred and a new page needs to be swapped in.
If the \texttt{working} buffer is full, a page fault on a memory access requires the VMM to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty and its changes have to be persisted to non-volatile memory (Figure~\ref{figure:mem-man}, Step 2). If the accessed page was previously modified and swapped out since the last power failure, the most recent version of the page is in {\tt \underline{shadow}} (Step 3), otherwise it has to be retrieved from {\tt \underline{private}} (Step 4). Finally, the page is copied to the volatile buffer (Step 5).

@@

To make the commit atomic, \sys commits dirty pages in two phases, as shown in Algorithm~\ref{algo:commit}. The first phase copies dirty pages from \texttt{working} to the non-volatile \texttt{\underline{shadow}} (Line 4). The second phase commits pages from \texttt{\underline{shadow}} to \texttt{\underline{private}} (Line 12).  If power fails during the first phase, the whole commit is aborted, and the execution restarts from the most recently committed point (i.e., from the beginning of a coalesced task). If power fails in the second
phase, the commit process safely resumes on reboot.  The second phase depends on some run-time metadata. The \texttt{\underline{committing}} bit indicates that a commit is in progress and is set before the first page is committed (Line 9) and cleared after the last page is committed (Line 16).  The \texttt{\underline{shadowCount}} records the number of dirty shadow pages to commit and the VMM clears the counter when commit completes (Line 14). The \texttt{\underline{commitIndex}} indexes the next page to be committed (Line 11) and the VMM clears the index at the end of the phase (Line 15).

\sys's commit is efficient because it maintains an index of dirty pages instead of iterating through all potentially dirty pages to check their state.
Another source of efficiency lies in the second phase of the commit:
% \sys's implementation is also efficient because
the VMM does not copy pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.

>>>>

@@

Resolving WAR (Write-After-Read) dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent when task are coalesced (see Fig.~\ref{figure:coalescedWar}). Therefore, \sys introduces its {\em page privatizer} (PP) which enables efficient task coalescing.

PP overview is given in Fig.~\ref{figure:mem-man}.
PP abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
\texttt{\underline{private}} holds the consistent version of pages after each commit (and on reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
The PP prohibits applications from directly accessing these buffers.
Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks.
When a coalesced task completes, the PP ensures that {\em dirty} (i.e., modified) pages are committed to FRAM and visible to subsequent tasks.
If a power failure interrupts a task, all temporary modifications to pages in \texttt{working} (and \texttt{\underline{shadow}}) occurred after the last commit are discarded in order to keep data consistency.
On a commit, the PP atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
% SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM. A shadow page stores the data from an updated page that needs to be committed back to non-volatile memory, but that exceeded the size of the volatile working memory.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/mem-man.pdf}
    \caption{Interaction between task, page privatizer and memory buffers upon \texttt{RP} and \texttt{WP}, and order of occurrence. Steps 2, 3, 4 and 5 are conditional.}
    \label{figure:mem-man}
\end{figure}

@@

After address translation (Fig.~\ref{figure:mem-man}, Step 6), a task accesses the protected variable's location in the volatile \texttt{working} buffer (Step 7). The PP keeps track of the page tags for the pages currently resident in the \texttt{working} buffer. When a task accesses a variable, it compares the variable's page tag to tags of the pages in {\tt working} (Algorithm~\ref{algo:rwar}, Line 2). If the accessed variable's page tag is not found in the page buffer, the operation incurs a {\em page fault} (Line 4). The byte is accessed in the page buffer at the index of the resident page and the variable's page offset (Lines 5--6).

@@

When accessing a protected variable with \texttt{RP} or \texttt{WP}, the PP first searches the variable's page in the \texttt{working} buffer (Fig.~\ref{figure:mem-man}, Step 1).
If the page is not found there, a page fault is incurred and a new page needs to be swapped in.
If the \texttt{working} buffer is full, a page fault on a memory access requires the PP to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty and its changes have to be persisted to non-volatile memory (Fig.~\ref{figure:mem-man}, Step 2). If the accessed page was previously modified and swapped out since the last power failure, the most recent version of the page is in {\tt \underline{shadow}} (Step 3), otherwise it has to be retrieved from {\tt \underline{private}} (Step 4). Finally, the page is copied to the volatile buffer (Step 5).

@@

To make the commit atomic, \sys commits dirty pages in two phases, as shown in Algorithm~\ref{algo:commit}. The first phase copies dirty pages from \texttt{working} to the non-volatile \texttt{\underline{shadow}} (Line 4). The second phase commits pages from \texttt{\underline{shadow}} to \texttt{\underline{private}} (Line 12).  If power fails during the first phase, the whole commit is aborted, and the execution restarts from the most recently committed point (i.e., from the beginning of a coalesced task). If power fails in the second
phase, the commit process safely resumes on reboot.  The second phase depends on some run-time metadata. The \texttt{\underline{committing}} bit indicates that a commit is in progress and is set before the first page is committed (Line 9) and cleared after the last page is committed (Line 16).  The \texttt{\underline{shadowCount}} records the number of dirty shadow pages to commit and the PP clears the counter when commit completes (Line 14). The \texttt{\underline{commitIndex}} indexes the next page to be committed (Line 11) and the PP clears the index at the end of the phase (Line 15).

\sys's commit is efficient because it maintains an index of dirty pages instead of iterating through all potentially dirty pages to check their state.
Another source of efficiency lies in the second phase of the commit:
% \sys's implementation is also efficient because
the PP does not copy pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.

### 2018-08-07 paper/asplos19/systemdescription.tex
change memory manager to page privatizer

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/system-overview.pdf}
	\caption{\sys top-level view. ATS: adaptive task scheduler, VMM: virtual memory manager, PM: physical memory.}
	\label{fig:system_overview}
\end{figure}

@@

\paragraph{Programming and Execution Model.}
To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate task-shared data. Then, they compile and link the code against \sys's runtime, producing a \sys-enabled binary. The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions. Facilitating efficient task adaptation requires dynamic memory protection which \sys's \emph{virtual memory manager} handles through a process we call \emph{dynamic page privatization}.

@@

\paragraph{Virtual Memory Manager.}
\sys can coalesce tasks thanks to its virtual memory manager (VMM). VMM privatizes memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory. VMM keeps all page modifications in non-volatile memory on a coalesced task transition. Section~\ref{sec:memory_virtulaization} outlines VMM.

>>>>

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/system-overview.pdf}
	\caption{\sys top-level view. ATS: adaptive task scheduler, PP: Page Privatizer, PM: physical memory.}
	\label{fig:system_overview}
\end{figure}

@@

\paragraph{Programming and Execution Model.}
To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate task-shared data. Then, they compile and link the code against \sys's runtime, producing a \sys-enabled binary. The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions. Facilitating efficient task adaptation requires dynamic memory protection which is what \sys's \emph{page privatizer} does.

@@

\paragraph{Page Privatizer.}
\sys can coalesce tasks thanks to its pager privatization (PP) technique. PP privatizes memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory. PP keeps all page modifications in non-volatile memory on a coalesced task transition. Section~\ref{sec:memory_virtulaization} outlines the PP mechanism.
